### 2021-10-08 paper.Rmd
working on results plot

```{r setup, include = FALSE}
library("papaja")
library("anticlust")
library("knitr")
options(digits = 2)
knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE)

@@

To optimize the three anticlustering criteria, a local-maximum search was applied that systematically swaps elements between groups to gradually improve the objective function [Method LCW in @weitz1998]. Building on an initial random assignment of elements to groups, elements are swapped between groups in such a way that each swap improves the objective criterion by the largest amount that is possible. That is, for each element, all possible swaps are simulated and the resulting objective values are stored. After each exchange has been tested, the one exchange is realized that improves the objective the most. No swap is realized if no swap improves the criterion. The exchange procedure is repeated for each element. Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements. The search stops as soon as an iteration through the entire set no longer yields any improvement, i.e., as soon as a local maximum has been found. To obtain better results, this local maximum search may be initialized multiple times [@spath1986]. This simulation employed five repetitions of the local maximum search.

@@

>>>>

```{r setup, include = FALSE}
library("papaja")
library("anticlust")
library("knitr")
options(digits = 2)
knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE, dev = "cairo_pdf")

@@

To optimize the three anticlustering criteria, a local-maximum search was applied that systematically swaps elements between groups to gradually improve the objective function [Method LCW in @weitz1998]. Building on an initial random assignment of elements to groups, elements are swapped between groups in such a way that each swap improves the objective criterion by the largest amount that is possible. That is, for each element, all possible swaps are simulated and the resulting objective values are stored. After each exchange has been tested, the one exchange is realized that improves the objective the most. No swap is realized if no swap improves the criterion. The exchange procedure is repeated for each element. Once the last element has been processed, the procedure restarts at the beginning and once again iterates through all elements. The search stops as soon as an iteration through the entire set no longer yields any improvement, i.e., as soon as a local maximum has been found. To obtain better results, this local maximum search may be initialized multiple times [@spath1986]. The simulation employed five repetitions of the local maximum search.

@@

```{r, fig.cap = "Displays how well maximizing the k-plus criterion approximates the k-means criterion that is found via direct maximization of the k-means criterion.", fig.height = 5, fig.width = 5}

# Load required packages
library(dplyr)
library(ggplot2)
library(tidyr)

## Analyze data for K = 2 and K = 3 and K = 4
simulation_results <- list()
for (K in 2:4) {
  filename <- paste0("./Simulation_Study/results-K", K, "-objectives-raw.csv")
  df <- read.csv(filename, sep = ";", stringsAsFactors = FALSE)
  df$K <- K
  simulation_results[[paste0("K-", K)]] <- df
}

df <- do.call(rbind, simulation_results)
rownames(df) <- NULL

# Make long format
ldf <- pivot_longer(
  df,
  cols = paste0(c("kvar", "kmeans", "means", "sd"), "_obj"),
  names_to = "Objective",
  names_pattern = "(.*)_obj"
)

# Check how well k-plus approximates k-means
rel <- ldf %>% 
  group_by(method, Objective, N, K) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(Objective == "kmeans", method %in% c("kplus", "k-means-exchange")) %>% 
  pivot_wider(names_from = method, values_from = Mean) %>% 
  group_by(N, K) %>% 
  summarise(rel = kplus / `k-means-exchange`)

rel %>% 
  ggplot(aes(x = N, y = rel, colour = factor(K), linetype = factor(K))) + 
  geom_line(size = 1) +
  geom_point(aes(colour = factor(K), shape = factor(K)), size = 3) +
  scale_x_continuous(breaks = seq(36, 300, 24)) + 
  theme_apa() +
  theme(
    legend.position = "top",
    legend.title = element_text()
  ) +
  scale_colour_grey(name = "K") +
  guides(
    linetype = guide_legend(title = "K"),
    shape = guide_legend(title = "K")
  ) + 
  labs(y = "k-plus approximation of k-means criterion")

# approximation at N > 100
# mean(filter(rel, N > 100)$rel) * 100

First, I investigated how well maximizing the k-plus criterion approximates the ordinary k-means criterion at the same time. Figure 2 illustrates the relative values of $\mathit{SSE}$ that were found when optimizing $\mathit{SSE_{kplus}}$ as compared to the direct optimization of $\mathit{SSE}$. Generally, optimizing $\mathit{SSE_{kplus}}$ yields an adequate approximation of $\mathit{SSE}$. However, the quality of this approximation depends on the group sizes of the partition problem: For larger $N$, k-plus anticlustering very closely follows k-means anticlustering's performance with regard to optimizing $\mathit{SSE}$. For $N > 100$, $\mathit{SSE}$ was approximated to more than 99.99% when conducting k-plus anticlustering as compared to k-means anticlustering. The minimum approximation for N = 24 and K = 4 was ...
